---
title: "getGBLUPs"
author: "LucianoRogerio"
date: "2022-07-11"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Previous step

3. [Check prediction accuracy](CrossValidation.html): Evaluate prediction accuracy with cross-validation.

## Objective

**Current Step**

4. Genomic prediction: Predict genomic BLUPs (GEBV and GETGV) for all selection candidates using all available data.

Set-up

```{bash, eval = F}
cd /home/jj332_cas/marnin/NRCRI_2021GS/; 
export OMP_NUM_THREADS=56 # activate multithread OpenBLAS 
```

```{r, eval = F}
.libPaths(c("~/R/x86_64-pc-linux-gnu-library/4.1", .libPaths()))
library(tidyverse); library(magrittr); 
source(here::here("code","gsFunctions.R"))
A<-readRDS(file=here::here("output","Kinship_A_NRCRI_2022Jul10.rds"))

# BLUPs from the 2 stage procedure
# (stage 1 of 2) 
blups<-readRDS(file=here::here("output","NRCRI_blupsForModelTraining_twostage_asreml_2022Jul09.rds")) %>% 
  select(Trait,blups) %>% 
  unnest(blups) %>% 
  select(-std.error) %>% 
  filter(GID %in% rownames(A)) %>% 
  nest(TrainingData=-Trait)
```

## Prediction

`runGenomicPredictions()`

cbsulm (112 cores; 512GB)

Model A

```{r, eval = F}
options(future.globals.maxSize= 1500*1024^2)
predModelA<-runGenomicPredictions(blups,modelType="A",grms=list(A=A),gid="GID",ncores=14)
saveRDS(predModelA,file = here::here("output","genomicPredictions_ModelA_twostage_NRCRI_2022Jul10.rds"))
```

Model ADE

```{r, eval = F}
D<-readRDS(file=here::here("output","Kinship_D_NRCRI_2022Jul10.rds"))
AD<-readRDS(file=here::here("output","Kinship_AD_NRCRI_2022Jul10.rds"))
options(future.globals.maxSize= 3000*1024^2)
predModelADE<-runGenomicPredictions(blups,modelType="ADE",grms=list(A=A,D=D,AD=AD),gid="GID",ncores=14)
saveRDS(predModelADE,file = here::here("output","genomicPredictions_ModelADE_twostage_NRCRI_2022Jul10.rds"))
```

## Write GEBV/GETV to disk

```{r}
rm(list=ls()); gc()


library(tidyverse); library(magrittr);
select <- dplyr::select
mutate <- dplyr::mutate
predModelA<-readRDS(file = here::here("output","genomicPredictions_ModelA_twostage_NRCRI_2022Jul10.rds"))
predModelADE<-readRDS(file = here::here("output","genomicPredictions_ModelADE_twostage_NRCRI_2022Jul10.rds"))
traits<-c("CGM", "CGMS1", "CGMS2", "MCMDS",
          "DMC", "DMCsg", "DRY",
          "IVig1-7", "PLTHT", "BRNHT1", "StC", "HI",
          "logDRYLD", "logDRYsgLD", "logFYLD", "logTOPYLD")

unique_gids<-predModelA %>%
  dplyr::select(genomicPredOut) %>%
  unnest(genomicPredOut) %>%
  dplyr::select(-varcomps) %>%
  unnest(gblups) %$%
  GID %>%
  unique

c1a<-unique_gids %>%
  grep("c1a",.,value = T,ignore.case = T) %>%
  union(.,unique_gids %>%
          grep("^F",.,value = T,ignore.case = T) %>%
          grep("c1b",.,value = T,ignore.case = T,invert = T))
c1b<-unique_gids%>% grep("c1b",.,value = T,ignore.case = T)
c2a<-unique_gids %>%
  grep("C2a",.,value = T,ignore.case = T) %>%
  grep("NR17",.,value = T,ignore.case = T)
c2b<-unique_gids %>%
  grep("C2b",.,value = T,ignore.case = T) %>%
  .[!. %in% c(c1a,c1b,c2a)]
c3a<-unique_gids %>%
  grep("C3a",.,value = T,ignore.case = T) %>%
  .[!. %in% c(c1a,c1b,c2a,c2b)]
c3b<-unique_gids %>%
  grep("NR20C3",.,value = T,ignore.case = T) %>%
  .[!. %in% c(c1a,c1b,c2a,c2b,c3a)]
c4a<-unique_gids %>%
  grep("NR21C4a",.,value = T,ignore.case = T) %>%
  .[!. %in% c(c1a,c1b,c2a,c2b,c3a,c3b)]
nrTP<-setdiff(unique_gids,unique(c(c1a,c1b,c2a,c2b,c3a,c3b,c4a)))

## Format and write GEBV
predModelA %>% 
  dplyr::select(Trait,genomicPredOut) %>% 
  unnest(genomicPredOut) %>% 
  dplyr::select(-varcomps) %>% 
  unnest(gblups) %>% 
  dplyr::select(-GETGV,-contains("PEV")) %>%
  spread(Trait,GEBV) %>% 
  mutate(Group=case_when(GID %in% nrTP ~ "nrTP",
                         GID %in% c1a ~ "C1a",
                         GID %in% c1b ~ "C1b",
                         GID %in% c2a ~ "C2a",
                         GID %in% c2b ~ "C2b",
                         GID %in% c3a ~ "C3a",
                         GID %in% c3b ~ "C3b",
                         GID %in% c4a ~ "C4a")) %>% 
  dplyr::select(Group,GID,any_of(traits)) %>% 
  arrange(desc(Group)) %>% 
  write.csv(., file = here::here("output","GEBV_NRCRI_ModelA_2022Jul10.csv"), row.names = F)

## Format and write GETGV
predModelADE %>% 
  dplyr::select(Trait,genomicPredOut) %>% 
  unnest(genomicPredOut) %>% 
  dplyr::select(-varcomps) %>% 
  unnest(gblups) %>% 
  dplyr::select(GID,Trait,GETGV) %>% 
  spread(Trait,GETGV) %>% 
  mutate(Group=case_when(GID %in% nrTP ~ "nrTP",
                         GID %in% c1a ~ "C1a",
                         GID %in% c1b ~ "C1b",
                         GID %in% c2a ~ "C2a",
                         GID %in% c2b ~ "C2b",
                         GID %in% c3a ~ "C3a",
                         GID %in% c3b ~ "C3b",
                         GID %in% c4a ~ "C4a")) %>% 
  dplyr::select(Group,GID,any_of(traits)) %>% 
  arrange(desc(Group)) %>% 
  write.csv(., file = here::here("output","GETGV_NRCRI_ModelADE_2022Jul10.csv"), row.names = F)

### Make a unified "tidy" long-form: 
predModelA %>% 
  dplyr::select(Trait,genomicPredOut) %>% 
  unnest(genomicPredOut) %>% 
  dplyr::select(-varcomps) %>% 
  unnest(gblups) %>% 
  dplyr::select(-GETGV) %>% 
  full_join(predModelADE %>% 
              dplyr::select(Trait,genomicPredOut) %>% 
              unnest(genomicPredOut) %>% 
              dplyr::select(-varcomps) %>% 
              unnest(gblups) %>% 
              rename(GEBV_modelADE=GEBV,
                     PEV_modelADE=PEVa)) %>% 
  mutate(Group=case_when(GID %in% nrTP ~ "nrTP",
                         GID %in% c1a ~ "C1a",
                         GID %in% c1b ~ "C1b",
                         GID %in% c2a ~ "C2a",
                         GID %in% c2b ~ "C2b",
                         GID %in% c3a ~ "C3a",
                         GID %in% c3b ~ "C3b",
                         GID %in% c4a ~ "C4a")) %>% 
  relocate(Group,.before = GID) %>% 
  write.csv(., file = here::here("output","genomicPredictions_NRCRI_2022Jul10.csv"), row.names = F)
```


## prepare Kinship matrix with correct names
```{r}
A <- readRDS(file = "output/Kinship_A_NRCRI_2022Jul10.rds")
Clones <- rownames(A)

ClonesCod <- read.table(file = here::here("data", "ClonesNamesNRCRI2022.csv"),
                        sep = ",", header = T)
rownames(ClonesCod) <- ClonesCod$GID
ClonesCod2 <- ClonesCod[Clones,]

rownames(A) <- colnames(A) <- ClonesCod2$GermplasmName
saveRDS(object = A, file = here::here("output", "Kinship_A_NRCRI_2022Aug16.rds"))
```

